---
title: "Snowball Search and global quantification of the references"
format:
    html:
        toc: true
        toc-depth: 2
        embed-resources: true
        code-fold: true
        code-summary: "Show the code"
    docx:
        toc: true
        toc-depth: 2
        number-sections: true
        highlight-style: github
    pdf:
        toc: true
        toc-depth: 2
        number-sections: true

---

# Setup
```{r}
#| label: setup
#|

library(bibtex)
library(openalexR)
library(writexl)
# library(dplyr)
# library(ggplot2)
# library(knitr)

kp <- bibtex::read.bib("./key-paper.bib")
dois <- sapply(
    kp,
    function(x) {
        x$doi
    }
)

```

# Searches

## OpenAlex

### Setup OpelAnex usage and do snowball serarch

```{r}
#| label: openalex_prep
#|

library(ggraph)
library(tidygraph)

key_works <- oa_fetch(
    entity = "works",
    doi = dois,
    verbose = FALSE
)

ids <- openalexR:::shorten_oaid(key_works$id)

citing_kp <- lapply(
    ids,
    function(id) {
        oa_fetch(
            entity = "works",
            output = "tibble",
            cites = id,
            verbose = FALSE
        )
    }
)
names(citing_kp) <- ids

cited_by_kp <- lapply(
    ids,
    function(id) {
        oa_fetch(
            entity = "works",
            output = "tibble",
            cited_by = id,
            verbose = FALSE
        )
    }
)
names(cited_by_kp) <- ids

if (file.exists("snowball.rds")) {
    snowball <- readRDS("snowball.rds")
} else {
    snowball <- oa_snowball(
        identifier = ids,
        verbose = FALSE
    )
    saveRDS(snowball, "snowball.rds")
}

flat_snow <- snowball2df(snowball) |>
    as_tibble()
```

### Save snowball as Excel file (`snowball_excel.xlsx`)

```{r}
#| label: openalex_excel
#| 

###
no_edges <- snowball$edges |>
    unlist() |>
    table() |>
    sort() |>
    as.data.frame() |>
    rename(
        no_connections = Freq,
        id = Var1
    )

###
no_referenced_works <- sapply(
    snowball$node["referenced_works"][[1]],
    length
)

no_referenced_works <- data.frame(
    id = snowball$node["id"],
    no_referenced_works = no_referenced_works
)

###
citations_per_year <- flat_snow |>
    select(
        id,
        publication_year,
        cited_by_count
    ) |>
    mutate(
        years_published = 2023 - publication_year
    ) |>
    mutate(
        avg_citations_per_year = cited_by_count / years_published
    ) |>
    arrange(desc(cited_by_count)) |>
    select(
        id,
        avg_citations_per_year
    ) |>
    rename(cited_global_per_year = avg_citations_per_year)

###

export <- flat_snow |>
    select(
        id,
        publication_year,
        display_name,
        doi,
        cited_by_count,
        ab
    ) |>
    rename(
        cited_global = cited_by_count,
        title = display_name,
        abstract = ab
    )


export$author <- sapply(
    flat_snow$author,
    function(z) {
        paste(unlist(z["au_display_name"]), collapse = ", ")
    }
)


export$author_institute <- sapply(
    flat_snow$author,
    function(z) {
        paste(unlist(z["institution_display_name"]), collapse = ", ")
    }
)


export$institute_country <- sapply(
    flat_snow$author,
    function(z) {
        paste(unlist(z["institution_country_code"]), collapse = ", ")
    }
)


concepts <- lapply(
    flat_snow$concepts,
    function(x){
        x |> 
        select(level, display_name, score)
    }
)

export$concepts_l0  <- sapply(
    concepts,
    function(x) {
       x |>
       filter(level == 0) |>
       mutate(name = paste0(display_name, " (", round(score, digits = 3), ")")) |>
       select(name) |>
       unlist() |>
       paste0(collapse = ", ")
    }
)

export$concepts_l1  <- sapply(
    concepts,
    function(x) {
       x |>
       filter(level == 1) |>
       mutate(name = paste0(display_name, " (", round(score, digits = 3), ")")) |>
       select(name) |>
       unlist() |>
       paste0(collapse = ", ")
    }
)

export$concepts_l2  <- sapply(
    concepts,
    function(x) {
       x |>
       filter(level == 2) |>
       mutate(name = paste0(display_name, " (", round(score, digits = 3), ")")) |>
       select(name) |>
       unlist() |>
       paste0(collapse = ", ")
    }
)

export$concepts_l3  <- sapply(
    concepts,
    function(x) {
       x |>
       filter(level == 3) |>
       mutate(name = paste0(display_name, " (", round(score, digits = 3), ")")) |>
       select(name) |>
       unlist() |>
       paste0(collapse = ", ")
    }
)

export$concepts_l4  <- sapply(
    concepts,
    function(x) {
       x |>
       filter(level == 4) |>
       mutate(name = paste0(display_name, " (", round(score, digits = 3), ")")) |>
       select(name) |>
       unlist() |>
       paste0(collapse = ", ")
    }
)

export$concepts_l5  <- sapply(
    concepts,
    function(x) {
       x |>
       filter(level == 5) |>
       mutate(name = paste0(display_name, " (", round(score, digits = 3), ")")) |>
       select(name) |>
       unlist() |>
       paste0(collapse = ", ")
    }
)


xlsx <- export |>
    full_join(no_edges, by = "id") |>
    full_join(no_referenced_works, by = "id") |>
    full_join(citations_per_year, by = "id") |>
    dplyr::relocate(author, .after = id) |>
    arrange(desc(cited_global))

la <- xlsx$abstract |>
    nchar() >= 3000

xlsx$abstract[la] <- substr(xlsx$abstract[la], 1, 3000)

xlsx |>
    dplyr::relocate(cited_global_per_year, .after = cited_global) |>
    dplyr::relocate(no_referenced_works, .after = doi) |>
    dplyr::relocate(no_connections, .before = abstract) |>
    dplyr::relocate(concepts_l0, .before = abstract) |>
    dplyr::relocate(concepts_l1, .before = abstract) |>
    dplyr::relocate(concepts_l2, .before = abstract) |>
    dplyr::relocate(concepts_l3, .before = abstract) |>
    dplyr::relocate(concepts_l4, .before = abstract) |>
    dplyr::relocate(concepts_l5, .before = abstract) |>
    dplyr::relocate(author_institute, .before = abstract) |>
    dplyr::relocate(institute_country, .before = abstract) |>
    writexl::write_xlsx(file.path(".", "data", "snowball_excel.xlsx"))
```

### Graph of links between references

```{r}
#| label: openalex_graph
#|

snowball$nodes$cited_by_count_by_year <- snowball$nodes$cited_by_count / (2024 - snowball$nodes$publication_year)

### Size cited_by_count_by_year
p_cby <- snowball |>
    as_tbl_graph() |>
    ggraph(graph = , layout = "stress") +
    geom_edge_link(aes(alpha = after_stat(index)), show.legend = FALSE) +
    geom_node_point(aes(fill = oa_input, size = cited_by_count_by_year), shape = 21, color = "white") +
    geom_node_label(aes(filter = oa_input, label = id), nudge_y = 0.2, size = 3) +
    scale_edge_width(range = c(0.1, 1.5), guide = "none") +
    scale_size(range = c(3, 10), guide = "none") +
    scale_fill_manual(values = c("#a3ad62", "#d46780"), na.value = "grey", name = "") +
    theme_graph() +
    theme(
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent", colour = NA),
        legend.position = "bottom"
    ) +
    guides(fill = "none") +
    ggtitle("Cited by average count per year")

ggsave("cited_by_count_by_year.pdf", plot = p_cby, device = cairo_pdf, width = 20, height = 15)
ggsave("cited_by_count_by_year.png", plot = p_cby, width = 20, height = 15, bg = "white")

### Size cited_by_count
p_cb <- snowball |>
    as_tbl_graph() |>
    ggraph(graph = , layout = "stress") +
    geom_edge_link(aes(alpha = after_stat(index)), show.legend = FALSE) +
    geom_node_point(aes(fill = oa_input, size = cited_by_count), shape = 21, color = "white") +
    geom_node_label(aes(filter = oa_input, label = id), nudge_y = 0.2, size = 3) +
    scale_edge_width(range = c(0.1, 1.5), guide = "none") +
    scale_size(range = c(3, 10), guide = "none") +
    scale_fill_manual(values = c("#a3ad62", "#d46780"), na.value = "grey", name = "") +
    theme_graph() +
    theme(
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent", colour = NA),
        legend.position = "bottom"
    ) +
    guides(fill = "none") +
    ggtitle("Cited by count")

ggsave("cited_by_count.pdf", plot = p_cb, device = cairo_pdf, width = 20, height = 15)
ggsave("cited_by_count.png", plot = p_cb, width = 20, height = 15, bg = "white")

```

### Identification of references with more than one edge

This is the number of connections (`connection_count)`of the paper (`id`)

```{r}
#| label: openalex_edgecount
#|

mult_edge <- flat_snow |>
    select(id, connection_count) |>
    filter(connection_count > 1) |>
    arrange(desc(connection_count))

links <- flat_snow |>
    filter(id %in% mult_edge$id)

links |>
    select(id, display_name, publication_year, doi, connection_count) |>
    arrange(desc(connection_count)) |>
    knitr::kable()
```


### Identification of Concepts

OpenAlex assigns all works concepts. The concepts are in hirarchical order, ranging from 0 to 3. The higher the number, the more specific the concept. The concepts are assigned to the paper (`id`)


#### Level 0
```{r}
#| label: l0
#| 

x <- lapply(
    flat_snow[["concepts"]],
    FUN = function(x) {
        x[["display_name"]][x[["level"]] == 0]
    }
) |>
    unlist() |>
    table() |>
    as.data.frame() |> 
    rename(
        l0_concept = Var1,
        count = Freq
    ) |>
    arrange(desc(count))

writexl::write_xlsx(x, file.path(".", "data", "concepts_l0.xlsx"))

knitr::kable(x)
```

#### Level 1
```{r}
#| label: l1
#| 

x <- lapply(
    flat_snow[["concepts"]],
    function(x) {
        x[["display_name"]][x[["level"]] == 1]
    }
) |>
    unlist() |>
    table() |>
    as.data.frame() |> 
    rename(
        l1_concept = Var1,
        count = Freq
    ) |>
    arrange(desc(count))
    
writexl::write_xlsx(x, file.path(".", "data", "concepts_l1.xlsx"))

knitr::kable(x)
```

#### Level 2

```{r}
#| label: l2
#| 

x <- lapply(
    flat_snow[["concepts"]],
    function(x) {
        x[["display_name"]][x[["level"]] == 2]
    }
) |>
    unlist() |>
    table() |>
    as.data.frame() |> 
    rename(
        l2_concept = Var1,
        count = Freq
    ) |>
    arrange(desc(count))
    
writexl::write_xlsx(x, file.path(".", "data", "concepts_l2.xlsx"))

knitr::kable(x)
```

#### Level 3

```{r}
#| label: l3
#| 

x <- lapply(
    flat_snow[["concepts"]],
    function(x) {
        x[["display_name"]][x[["level"]] == 3]
    }
) |>
    unlist() |>
    table() |>
    as.data.frame() |> 
    rename(
        l3_concept = Var1,
        count = Freq
    ) |>
    arrange(desc(count))
    
writexl::write_xlsx(x, file.path(".", "data", "concepts_l3.xlsx"))

knitr::kable(x)
```

#### Level 4

```{r}
#| label: l4
#| 

x <- lapply(
    flat_snow[["concepts"]],
    function(x) {
        x[["display_name"]][x[["level"]] == 4]
    }
) |>
    unlist() |>
    table() |>
    as.data.frame() |> 
    rename(
        l4_concept = Var1,
        count = Freq
    ) |>
    arrange(desc(count))
    
writexl::write_xlsx(x, file.path(".", "data", "concepts_l4.xlsx"))
    
knitr::kable(x)
```

#### Level 5

```{r}
#| label: l5
#| 

x <- lapply(
    flat_snow[["concepts"]],
    function(x) {
        x[["display_name"]][x[["level"]] == 5]
    }
) |>
    unlist() |>
    table() |>
    as.data.frame() |> 
    rename(
        l5_concept = Var1,
        count = Freq
    ) |>
    arrange(desc(count))
    
writexl::write_xlsx(x, file.path(".", "data", "concepts_l5.xlsx"))

knitr::kable(x)
```

